WIDS 2024 Final Project

### 2024-01-09 Code/R FIle.qmd
updated x axis to be every 4 years

##### Min Wage by state and year (include fed (facet or use by decades where necessary)
```{r}
state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <=2020 ) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "OK" ~ "OK",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
state_minimum_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(state_minimum_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = state_min_wage_2020_adj, color = map_color),
            linewidth = .75) +
  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) + # federal minimum wage
  theme_minimal() +
  labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
       subtitle = "For states with an adjusted effective minimum wage above the federal level",
       x = "Year",
       y = "Minimum wage (2020 USD)")

@@

##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
```{r}
effective_minimum_wage_data <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA",
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "CT" ~ "CT",
      state_abbreviation == "HI" ~ "HI",
      .default = NA
    )
  ) 
effective_minimum_wage_data |> 
  ggplot(aes(x = year)) +
  geom_line(aes(y = effective_min_wage_2020_adj, group = state_abbreviation),
            color = "grey75", alpha =0.6) +
  geom_line(data = subset(effective_minimum_wage_data, !is.na(map_color)),
            aes(y = effective_min_wage_2020_adj, color = map_color),
            linewidth = .75) +
  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
  theme_minimal() +
  labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
       subtitle = "For states with an adjusted effective minimum wage above the federal level",
       x = "Year",
       y = "Effective minimum wage (2020 USD)")

@@

##### percent union and percent collective bargaining
```{r}
per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <=2020 ) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "OK" ~ "OK",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
per_union_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = percent_union, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(per_union_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = percent_union, color = map_color),
            linewidth = .75) +
  theme_minimal() +
  labs(title = "States by % of labor market covered by union",
       x = "Year",
       y = "% of labor market in union")

@@

```{r}
per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <=2020 ) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "HI" ~ "HI",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
per_collective_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(per_collective_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = percent_collective_bargain, color = map_color),
            linewidth = .75) +
  theme_minimal() +
  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
       x = "Year",
       y = "% of labor market in union")

>>>>

##### Min Wage by state and year (include fed (facet or use by decades where necessary)
```{r}
state_minimum_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <=2020 & year >=1984) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "OK" ~ "OK",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
state_minimum_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = state_min_wage_2020_adj, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(state_minimum_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = state_min_wage_2020_adj, color = map_color),
            linewidth = .75) +
  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) + # federal minimum wage
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(state_minimum_wage_data$year), max(state_minimum_wage_data$year), by = 4)) + 
  labs(title = "State Minimum Wage by Year (Adjusted for 2020 $)",
       subtitle = "For states with an adjusted effective minimum wage above the federal level",
       x = "Year",
       y = "Minimum wage (2020 USD)")

@@

##### Adjusted Min Wage by state and year (include adjusted fed (facet or use by decades where necessary)
```{r}
effective_minimum_wage_data <- full_state_data |> 
  filter(!is.na(regions)) |> 
  filter(effective_min_wage_2020_adj != fed_min_wage_2020_adj) |> 
  filter(year <=2020 & year >=1984) |>
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA",
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "CT" ~ "CT",
      state_abbreviation == "HI" ~ "HI",
      .default = NA
    )
  ) 
effective_minimum_wage_data |> 
  ggplot(aes(x = year)) +
  geom_line(aes(y = effective_min_wage_2020_adj, group = state_abbreviation),
            color = "grey75", alpha =0.6) +
  geom_line(data = subset(effective_minimum_wage_data, !is.na(map_color)),
            aes(y = effective_min_wage_2020_adj, color = map_color),
            linewidth = .75) +
  geom_line(aes(y = fed_min_wage), color = "red", linewidth = .75) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(effective_minimum_wage_data$year), max(effective_minimum_wage_data$year), by = 4)) + 
  labs(title = "State Effective Minimum Wage by Year (Adjusted for 2020 $)",
       subtitle = "For states with an adjusted effective minimum wage above the federal level",
       x = "Year",
       y = "Effective minimum wage (2020 USD)")

@@

##### percent union and percent collective bargaining
```{r}
per_union_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "OK" ~ "OK",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
per_union_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = percent_union, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(per_union_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = percent_union, color = map_color),
            linewidth = .75) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(per_union_wage_data$year), max(per_union_wage_data$year), by = 4)) + 
  labs(title = "States by % of labor market covered by union",
       x = "Year",
       y = "% of labor market in union")

@@

```{r}
per_collective_wage_data <- full_state_data |> # This object will be used to create the graph and also in the data subset
  filter(!is.na(regions)) |> #This just removes DC
  filter(year <= 2020 & year >= 1984) |>  #wage data only exists before 2020
  mutate(
    map_color = case_when(
      state_abbreviation == "CA" ~ "CA", # States we would like to highlight. first "" is the state abb, second "" is what will appear in legend
      state_abbreviation == "AK" ~ "AK",
      state_abbreviation == "HI" ~ "HI",
      state_abbreviation == "FL" ~ "FL",
      .default = NA
    )
  ) 
per_collective_wage_data |> # use created table here
  ggplot(aes(x = year)) +
  geom_line(aes(y = percent_collective_bargain, group = state_abbreviation),
            color = "grey75", alpha =0.6) + # using group means the groupings wont appear in legend
  geom_line(data = subset(per_collective_wage_data, # use created object here
                          !is.na(map_color)), #subset used to keep out non highlighted states
            aes(y = percent_collective_bargain, color = map_color),
            linewidth = .75) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(per_collective_wage_data$year), max(per_collective_wage_data$year), by = 4)) + 
  labs(title = "States by % of labor market covered by collective Bargaining Agreement",
       x = "Year",
       y = "% of labor market in union")
